Katana: Towards Patching as a Part of the ELF Toolchain

### 2010-03-18 paper1/hotpatch/poformat.tex
removed numbered subsection headings, as required by style for this publication

\section{Patch Object Format}
\label{sec:poformat}
\subsection{Reasons and Needs}
We have developed a Patch Object format with the following properties
\begin{itemize}
 \item a valid ELF file
 \item information about types and functions requiring patches contained
   as DWARF information
 \item type transformations specified using a language defined by the
   Dwarf standard
\end{itemize}
Through the use of existing standards and well-structured ELF files
utilizing a simple expression language for data patching, we aim to
create patches which are easily examined (or modified) with existing
tools. This easy compatibility with the existing binary tools and
standards brings us to a very good point: why should patching not be a
part of the ABI and of the standard toolchain? This does not
necessarily have to be the precise format we use for Katana, any such
format that would become a standard, whether an actual standard or a
de-facto standard should be well vetted by the community, but we argue
that something like this should be included in the standard object
types, along with relocatable objects, executable objects, shared
libraries, and core dumps. Consider the situation. Relocatable objects
containing new code and data which may be inserted at runtime are
nothing new. This is the entire premise of the dynamic
library. User-written functions which may have to run upon this code
injection (in the case of patching data where the desired actions
cannot be determined automatically) already exist as the
\texttt{.init} and \texttt{.fini} sections. Due to this similarity
between some of the functionality needed by patching and the
functionality offered by dynamic
libraries, some previous systems have performed patching by creating
patches as dynamic libraries which contain not only the code and data
to be patched but the mechanism to perform the patching \cite{ginseng}
\cite{polus}. We argue that
this is an unnecessary mixing of data and logic, and that a patch
which contains merely the information necessary to fix a running
process and not the code to do so is more desirable. The code to apply
the patch should live in one place on any given system, as most other
executable content does. We do not embed Emacs within our text files,
after all. Dynamic libraries and other relocatable, linkable objects
do not contain code and data intended to overwrite data in an existing
executable or process. Consider, however, that redefining certain
symbols is only a slight twist on ordinary linker behaviour. Ordinary
linker behaviour for global symbols is to fail if a symbol is defined
more than once. Ordinary behaviour for weak symbols is to use a global
definition if available or the weak symbol otherwise. When performing
dynamic linking, generally the first appropriate symbol encountered in
the chain of symbol tables is used. It is not a far difference to
define the linkage rule that the symbol definition from the most
recent patch takes precedence. Therefore, applying a patch consists of
the following steps
\begin{enumerate}
\item Injecting appropriate sections of the patch into memory. This
includes putting their contents into memory and performing relocations
on these sections (but not on the rest of the in-memory process) so
that they fit into their environment
\item Copying existing data to the appropriate regions of the newly
mapped-in patch
\item Performing relocation on the entire in-memory process such that the
symbols defined by the patch take precedence.
\end{enumerate}
These steps are all such fundamental operations that they should
become universally supported by the ABI and the toolchain.

@@

\subsection{Our Patch Object Format}
Our Patch Object (PO) format is an ELF-based format. Figure
\ref{fig:elfheaders} shows the sections contained in a simple
patch. \texttt{.text.new} and \texttt{.rodata.new} are of course the
new code and supporting

>>>>

\section{Patch Object Format}
\label{sec:poformat}
%\subsection{Reasons and Needs}
{\bf Reasons and Needs.}
We have developed a Patch Object format with the following properties
\begin{itemize}
 \item a valid ELF file
 \item information about types and functions requiring patches contained
   as DWARF information
 \item type transformations specified using a language defined by the
   Dwarf standard
\end{itemize}
Through the use of existing standards and well-structured ELF files
utilizing a simple expression language for data patching, we aim to
create patches which are easily examined (or modified) with existing
tools. This easy compatibility with the existing binary tools and
standards brings us to a very good point: why should patching not be a
part of the ABI and of the standard toolchain? This does not
necessarily have to be the precise format we use for Katana, any such
format that would become a standard, whether an actual standard or a
de-facto standard should be well vetted by the community, but we argue
that something like this should be included in the standard object
types, along with relocatable objects, executable objects, shared
libraries, and core dumps. Consider the situation. Relocatable objects
containing new code and data which may be inserted at runtime are
nothing new. This is the entire premise of the dynamic
library. User-written functions which may have to run upon this code
injection (in the case of patching data where the desired actions
cannot be determined automatically) already exist as the
\texttt{.init} and \texttt{.fini} sections. Due to this similarity
between some of the functionality needed by patching and the
functionality offered by dynamic
libraries, some previous systems have performed patching by creating
patches as dynamic libraries which contain not only the code and data
to be patched but the mechanism to perform the patching \cite{ginseng}
\cite{polus}. We argue that
this is an unnecessary mixing of data and logic, and that a patch
which contains merely the information necessary to fix a running
process and not the code to do so is more desirable. The code to apply
the patch should live in one place on any given system, as most other
executable content does. We do not embed Emacs within our text files,
after all. Dynamic libraries and other relocatable, linkable objects
do not contain code and data intended to overwrite data in an existing
executable or process. Consider, however, that redefining certain
symbols is only a slight twist on ordinary linker behaviour. Ordinary
linker behaviour for global symbols is to fail if a symbol is defined
more than once. Ordinary behaviour for weak symbols is to use a global
definition if available or the weak symbol otherwise. When performing
dynamic linking, generally the first appropriate symbol encountered in
the chain of symbol tables is used. It is not a far difference to
define the linkage rule that the symbol definition from the most
recent patch takes precedence. Therefore, applying a patch consists of
the following steps
\begin{enumerate}
\item Injecting appropriate sections of the patch into memory. This
includes putting their contents into memory and performing relocations
on these sections (but not on the rest of the in-memory process) so
that they fit into their environment
\item Copying existing data to the appropriate regions of the newly
mapped-in patch
\item Performing relocation on the entire in-memory process such that the
symbols defined by the patch take precedence.
\end{enumerate}
These steps are all such fundamental operations that they should
become universally supported by the ABI and the toolchain.

@@

%\subsection{Our Patch Object Format}
{\bf Our Patch Object Format.}
Our Patch Object (PO) format is an ELF-based format. Figure
\ref{fig:elfheaders} shows the sections contained in a simple
patch. \texttt{.text.new} and \texttt{.rodata.new} are of course the
new code and supporting
